---
title: "S6. bigMART Analysis"
author: "Lutfor Rahman"
date: "2023-11-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(srvyr)
bm<-readRDS("tom_finished.rds") #%>% filter(Demographic.FishyV3 == "Not so fishy")
names(bm)

test1<-bm %>% select(Income.Total)
```


# 2 Percentage and number of the population currently in a Negative Budget + change over time.
```{r}
#1 Percentage and number of the population currently in a Negative Budget + change over time.
##What % are in negative budget
bm %>% as_survey_design(weights = Weight.LCFS)%>% 
        group_by(Financial.Year,Demographic.Negbud)%>%
                  summarise(n=survey_total()) %>% 
                  mutate(per_n=n/sum(n))%>%
                  mutate(population=case_when (Financial.Year!="2020-21"~(n*1000/2),
                                               TRUE ~ (n*1000)
                                               )) %>%
        filter(Demographic.Negbud == "Negative.Budget") %>%
        select(-(contains("_se"))) %>%
        tibble()
```

# 2a Weekly surplus
```{r}
#weekly surplus and income by negative budget

bm %>% 
  as_survey_design(weights=Weight.LCFS) %>% 
  group_by(Financial.Year,Demographic.Negbud)%>%
  summarise(tmean_surplus = survey_mean(Surplus, na.rm=TRUE),
            median_surplus = survey_median(Surplus, na.rm=TRUE),
            mean_income = survey_mean(Income.Equivalised, na.rm=TRUE),
            median_income = survey_median(Income.Equivalised, na.rm=TRUE)) %>% 
   select(-(contains("_se"))) %>%
  tibble()
```

# 3 Percentage and number of the population currently on the brink (have less than £100 spare each month?) + change over time. 
```{r}
#3. Percentage and number of the population currently on the brink (have less than £100 spare each month?) + change over time. 

brink_bm<-bm %>% mutate(awayfromcrisis= case_when(Surplus<(-50)~"More_than_-£50",
                                   Surplus>=-50&Surplus<(-20)~"£50_inRed",
                                   Surplus>=-20&Surplus<(-10)~"£20_inRed", 
                                   Surplus>=-10&Surplus<0~"£10_inRed",
                                   Surplus>=0&Surplus<=10~"£10_away",
                                   Surplus>10&Surplus<=20~"£20_away",
                                   Surplus>20&Surplus<=50~"£50_away",
                                   TRUE~"More_than_£50"))


 brink_bm %>% as_survey_design(weights = Weight.LCFS)%>% 
        filter(Financial.Year=="2023-24") %>% 
        group_by(awayfromcrisis)%>%
        summarise(n=survey_total(),
                  population=n*1000/2,
                  Median_surplus=survey_median(Surplus, na.rm=TRUE),
                  Mean_surplus=survey_mean(Surplus,na.rm=TRUE)
        ) %>% 
        
        select(-(contains("_se"))) %>%
        arrange(Mean_surplus) %>%
        as.data.frame()
```


# 3a.Percentage and number of the population currently on the brink (have less than £100 spare each month?) + change over time. 
```{r}
brink_bm %>% as_survey_design(weights = Weight.LCFS)%>% 
        filter(Financial.Year=="2023-24") %>% 
        group_by(Demographic.Surplus.Bands.V1)%>%
        summarise(n=survey_total(),
                  population=n*1000,
                  Median_surplus=survey_median(Surplus, na.rm=TRUE),
                  Mean_surplus=survey_mean(Surplus, na.rm=TRUE)
        ) %>% 
        
        select(-(contains("_se"))) %>%
        arrange(Mean_surplus) %>%
        as.data.frame()
```

# 5. The percentage and number of people in a NB before we reduce LCFS respondent’s flexible spending to our MART data averages.
```{r}
#The percentage and number of people in a NB before we reduce LCFS respondent’s flexible spending to our MART data averages.

LCFS_Only_Negative_Budget<-bm %>% mutate(Expenditure.Fixed.Flexible.LCF=Expenditure.Fixed.Total+Expenditure.Flexible.Total.LCFS) %>% 
   mutate(Surplus.LCFS = (Income.Total - (Expenditure.Fixed.Total + Expenditure.Flexible.Total.LCFS  ))) %>% 
   mutate(Negbud.LCFS= case_when(Surplus.LCFS<=0 ~ 'Negative.Budget',
                           TRUE ~ 'Positive.Budget')) 


LCFS_Only_Negative_Budget %>% as_survey_design(weights = Weight.LCFS)%>% 
  group_by(Financial.Year,Negbud.LCFS)%>%
  summarise(n=survey_total()) %>% 
  mutate(per_n=n/sum(n))%>%
  mutate(population=case_when (Financial.Year!="2020-21"~(n*1000/2),
                               TRUE ~ (n*1000)
                               )) %>%
  filter(Negbud.LCFS == "Negative.Budget") %>%
  select(-(contains("_se"))) %>%
  tibble()
```



# 6. Demographics of those most affected by NB nationally 
```{r}
#Demographics of those most affected by NB nationally
variables_to_group_by <- c("Derived.Household.Type.LCFS", 
                           "Housing",
                           "Role.Type",
                           "Disabled", 
                           "UC",
                           "ethnicity.HRP")
# To understand negative budget by demographic groups 
results_negative_demographic <- map_dfr(variables_to_group_by, function(var) {
  bm %>% 
    as_survey_design(weights = Weight.LCFS) %>%
    filter(Financial.Year == "2023-24") %>% 
    group_by(!!sym(paste0("Demographic.", var)), Demographic.Negbud) %>%
    summarise(n = survey_total()) %>%
    mutate(per_n = n / sum(n)) %>%
    mutate(population = n * 1000 / 2) %>%
    filter(Demographic.Negbud == "Negative.Budget") %>%
    select(-(contains("_se"))) %>%
    tibble()
})

print(results_negative_demographic)
write.csv(results_negative_demographic, "results_negative_demographic.csv")

```

# 6. Demographics of those most affected by NB nationally and their surplus
```{r}
#Demographics of those most affected by NB nationally and their surplus

variables_to_group_by <- c("Derived.Household.Type.LCFS", 
                           "Housing",
                           "Role.Type",
                           "Disabled", 
                           "UC",
                           "ethnicity.HRP")

results_surplus_demographic <- map_dfr(variables_to_group_by, function(var) {
    bm %>% 
    as_survey_design(weights = Weight.LCFS) %>%
    filter(Financial.Year == "2023-24") %>% 
    group_by(!!sym(paste0("Demographic.", var)), Demographic.Negbud) %>%
    summarise(
      mean_surplus = survey_mean(Surplus, na.rm = TRUE),
      median_surplus = survey_median(Surplus, na.rm = TRUE)
                             ) %>%
    filter(Demographic.Negbud == "Negative.Budget") %>%
    select(-(contains("_se"))) %>%
    tibble()
})

print(results_surplus_demographic)

write.csv(results_surplus_demographic, "results_surplus_demographic.csv")

```


# 7 What are biggest driver of negative budget?
```{r}
#7 What are biggest driver of negative budget? To understand which proportion of income goes into expenditure type

variables_to_group_by <- c("Derived.Household.Type.LCFS", 
                           "Housing",
                           "Role.Type",
                           "Disabled", 
                           "UC",
                           "ethnicity.HRP")

driver<-bm %>% 
  mutate(housing_ratio = case_when(Expenditure.Housing <= 0 ~ NA,
                                   TRUE ~ Expenditure.Housing / Income.Total),
         rent_ratio = case_when(Expenditure.Rent <= 0 ~ NA,
                                   TRUE ~ Expenditure.Rent / Income.Total),
         mortgage_ratio = case_when(Expenditure.Mortgage <= 0 ~ NA,
                                    TRUE ~ Expenditure.Mortgage / Income.Total),
         transport_ratio = case_when(Expenditure.Transport <= 0 ~ NA,
                                     TRUE ~ Expenditure.Transport / Income.Total),
         utilities_ratio = case_when(Expenditure.Utilities.MART <= 0 ~ NA,
                                     TRUE ~ Expenditure.Utilities.MART / Income.Total),
         food_ratio = case_when(Expenditure.Food.House.Keeping.MART <= 0 ~ NA,
                                     TRUE ~ Expenditure.Food.House.Keeping.MART / Income.Total),
         essential_ratio = case_when(Expenditure.Total <= 0 ~ NA,
                                     TRUE ~ Expenditure.Total / Income.Total)
  )

test<-driver %>% select(housing_ratio, rent_ratio, mortgage_ratio, transport_ratio, utilities_ratio,
                        food_ratio, essential_ratio, Expenditure.Housing,Expenditure.Rent ,Income.Total, Expenditure.Total)

write.csv(test, "test.csv")


#This is to without demographic looking into expenditure ratio
result_driver_overall <-driver %>% 
                        as_survey_design(weights = Weight.LCFS) %>%
                            group_by(Financial.Year, Demographic.Negbud) %>%
                            summarise(
                              mean_housing_ratio = survey_mean(housing_ratio, na.rm = TRUE),
                              median_housing_ratio = survey_median(housing_ratio, na.rm = TRUE),
                              
                              mean_rent_ratio = survey_mean(rent_ratio, na.rm = TRUE),
                              median_rent_ratio = survey_median(rent_ratio, na.rm = TRUE),
                              
                              mean_mortgage_ratio = survey_mean(mortgage_ratio, na.rm = TRUE),
                              median_mortgage_ratio = survey_median(mortgage_ratio, na.rm = TRUE),
                  
                              
                              mean_trans_ratio = survey_mean(transport_ratio, na.rm = TRUE),
                              median_trans_ratio = survey_median(transport_ratio, na.rm = TRUE),
                              
                              mean_utilities_ratio = survey_mean(utilities_ratio, na.rm = TRUE),
                              median_utilities_ratio = survey_median(utilities_ratio, na.rm = TRUE),
                                          
                              mean_food_ratio = survey_mean(food_ratio, na.rm = TRUE),
                              median_food_ratio = survey_median(food_ratio, na.rm = TRUE),
                             
                              mean_essential_ratio = survey_mean(essential_ratio, na.rm = TRUE),
                              median_essential_ratio = survey_median(essential_ratio, na.rm = TRUE),
                             
                            ) %>%
                             select(-(contains("_se"))) %>%
                             pivot_longer(cols = starts_with("mean_") | starts_with("median_"),
                             names_to = c(".value", "measure"),
                             names_sep = "_") %>%
                            tibble()


print(result_driver_overall)
write.csv(result_driver_overall, "result_driver_overall.csv")
```


```{r}
result_driver_demo_utilities <- map_dfr(variables_to_group_by, function(var) {
  driver %>% 
as_survey_design(weights = Weight.LCFS) %>%
    filter(Demographic.Negbud=="Negative.Budget") %>% 
    group_by(Financial.Year, !!sym(paste0("Demographic.", var))) %>%
    summarise(
      mean_utilities_ratio = survey_mean(utilities_ratio, na.rm = TRUE),
      median_utilities_ratio = survey_median(utilities_ratio, na.rm = TRUE)
    ) %>%
    select(-(contains("_se"))) %>%
    tibble()
})

print(result_driver_demo_utilities)

write.csv(result_driver_demo_utilities, "result_driver_demo_utilities.csv")



result_driver_demo_housing <- map_dfr(variables_to_group_by, function(var) {
  driver %>% 
as_survey_design(weights = Weight.LCFS) %>%
    group_by(Financial.Year, !!sym(paste0("Demographic.", var))) %>%
    summarise(
      mean_housing_ratio = survey_mean(housing_ratio, na.rm = TRUE),
      median_housing_ratio = survey_median(housing_ratio, na.rm = TRUE)
    )  %>%
    select(-(contains("_se"))) %>%
    tibble()
})

print(result_driver_demo_housing)

write.csv(result_driver_demo_housing, "result_driver_demo_housing.csv")

result_driver_demo_food <- map_dfr(variables_to_group_by, function(var) {
  driver %>% 
as_survey_design(weights = Weight.LCFS) %>%
    group_by(Financial.Year, !!sym(paste0("Demographic.", var))) %>%
    summarise(
      mean_food_ratio = survey_mean(food_ratio, na.rm = TRUE),
      median_food_ratio = survey_median(food_ratio, na.rm = TRUE)
    )  %>%
    select(-(contains("_se"))) %>%
    tibble()
})

write.csv(result_driver_demo_food, "result_driver_demo_food.csv")
```


# 11. By the end of 2023, people will have built up £Xbn of debt because their wages (and/or benefits?) can’t cover their costs. 
```{r}
#By the end of 2023, people will have built up £Xbn of debt because their wages (and/or benefits?) can’t cover their costs. 
debt_23_24 <- bm %>% 
  as_survey_design(weights = Weight.LCFS) %>% 
  filter(Financial.Year == "2023-24") %>% 
  group_by(Financial.Year, Demographic.Negbud) %>%
  summarise(
    n = survey_total(),
    mean_surplus = survey_mean(Surplus, na.rm = TRUE, trim = 0.2),
    median_surplus = survey_median(Surplus, na.rm = TRUE)
  ) %>% 
  mutate(
    per_n = n / sum(n),
    total_debt_mean = mean_surplus * n*52.1*1000/2, #to get household number, n need to multiply 1000
    total_debt_median = median_surplus * n*52.1*1000/2
  ) %>%
  select(-(contains("_se"))) %>%
  tibble()

print(debt_23_24)
write.csv(debt_23_24, "debt_23_24.csv")
```


#8 What share of the UC standard allowance is taken up by housing + energy for people renting privately, in social housing & mortgage holders.

```{r}
#What share of the UC standard allowance is taken up by housing + energy for people renting privately, in social housing & mortgage holders
UC_analysis<-bm %>%
  filter(Financial.Year=="2023-24") %>% 
  mutate(
    UC_Standard_Single_U25 = 292.11*12/52.143,
    UC_Standard_Single_O25 = 368.74*12/52.143,
    UC_Standard_Couple_U25 = 458.51*12/52.143,
    UC_Standard_Couple_O25 = 578.82*12/52.143
  ) %>% 
  mutate(
    UC_Value = case_when(
      # If single and under 25 then single standard allowance for u25
      Demographic.Derived.Household.Type.LCFS == "Single Person" & Demographic.Age.Group.HRP %in% c("20 but under 25 yrs", "15 but under 20 yrs") ~ UC_Standard_Single_U25,
      # Else if single then single standard allowance over 25
      Demographic.Derived.Household.Type.LCFS == "Single Person" ~ UC_Standard_Single_O25,
            # If single and under 25 then single standard allowance for u25
      Demographic.Derived.Household.Type.LCFS == "Single Person With Dependent Children" & Demographic.Age.Group.HRP %in% c("20 but under 25 yrs", "15 but under 20 yrs") ~ UC_Standard_Single_U25,
      # Else if single then single standard allowance over 25
      Demographic.Derived.Household.Type.LCFS == "Single Person With Dependent Children" ~ UC_Standard_Single_O25,

      # If couple and under 25 then couple standard allowance for u25
      Demographic.Derived.Household.Type.LCFS == "Couple" & Demographic.Age.Group.HRP %in% c("20 but under 25 yrs", "15 but under 20 yrs") ~ UC_Standard_Couple_U25,
      # Else if couple then couple standard allowance over 25
      Demographic.Derived.Household.Type.LCFS == "Couple" ~ UC_Standard_Couple_O25,
      
            # If couple and under 25 then couple standard allowance for u25
      Demographic.Derived.Household.Type.LCFS == "Couple With Dependent Children" & Demographic.Age.Group.HRP %in% c("20 but under 25 yrs", "15 but under 20 yrs") ~ UC_Standard_Couple_U25,
      # Else if couple then couple standard allowance over 25
      Demographic.Derived.Household.Type.LCFS == "Couple With Dependent Children" ~ UC_Standard_Couple_O25,

      # If "Adults only" and under 25 then single standard allowance for u25
      Demographic.Derived.Household.Type.LCFS == "Adults only" & Demographic.Age.Group.HRP %in% c("20 but under 25 yrs", "15 but under 20 yrs") ~ UC_Standard_Single_U25,
      # Else if "Adults only" then single standard allowance over 25
      Demographic.Derived.Household.Type.LCFS == "Adults only" ~ UC_Standard_Single_O25,

      # If "Other adults and Dependent children" and under 25 then couple standard allowance for u25
      Demographic.Derived.Household.Type.LCFS == "Other adults and Dependent children" & Demographic.Age.Group.HRP %in% c("20 but under 25 yrs", "15 but under 20 yrs") ~ UC_Standard_Couple_U25,
      # Else is "Other adults and Dependent children" then couple standard allowance over 25
      Demographic.Derived.Household.Type.LCFS == "Other adults and Dependent children" ~ UC_Standard_Couple_O25,
      TRUE ~ UC_Standard_Couple_O25
    )
  ) %>%
  mutate(UC_Value = case_when(
    UC_Value > Income.UC ~ Income.UC,
    TRUE ~ UC_Value
  )) %>% 
  filter(Income.UC>0) %>% 
  filter(Income.Gross.Wages.Respondent<=0 & Income.Gross.Wages.Partner<=0)

##case when expenditure is zero
UC_driver <- UC_analysis %>%
  mutate(housing_ratio = case_when(Expenditure.Housing <= 0 ~ NA,
                                   TRUE ~ Expenditure.Housing / UC_Value),
         rent_ratio = case_when(Expenditure.Rent <= 0 ~ NA,
                                TRUE ~ Expenditure.Rent / UC_Value),
         transport_ratio = case_when(Expenditure.Transport <= 0 ~ NA,
                                     TRUE ~ Expenditure.Transport / UC_Value),
         utilities_ratio = case_when(Expenditure.Utilities.MART <= 0 ~ NA,
                                     TRUE ~ Expenditure.Utilities.MART / UC_Value),
         food_ratio = case_when(Expenditure.Food.House.Keeping.MART <= 0 ~ NA,
                                TRUE ~ Expenditure.Food.House.Keeping.MART / UC_Value),
         essential_ratio = case_when(Expenditure.Total <= 0 ~ NA,
                                     TRUE ~ Expenditure.Total / UC_Value)
  ) %>% 
  mutate(Expenditure.Housing.Utilities=Expenditure.Housing+Expenditure.Utilities.MART) %>% 
  mutate(Housing.Utilities_ratio = case_when(Expenditure.Housing.Utilities <= 0 ~ NA,
                                   TRUE ~ Expenditure.Housing.Utilities / UC_Value))

##

UC_driver %>% 
as_survey_design(weights = Weight.LCFS) %>%
   # group_by(Financial.Year)%>%
    summarise(
      mean_housing_ratio = survey_mean(housing_ratio, na.rm = TRUE),
      median_housing_ratio = survey_median(housing_ratio, na.rm = TRUE),
      
      mean_rent_ratio = survey_mean(rent_ratio, na.rm = TRUE),
      median_rent_ratio = survey_median(rent_ratio, na.rm = TRUE),
      
      mean_transport_ratio = survey_mean(transport_ratio, na.rm = TRUE),
      median_transport_ratio = survey_median(transport_ratio, na.rm = TRUE),
      
      mean_utilities_ratio = survey_mean(utilities_ratio, na.rm = TRUE),
      median_utilities_ratio = survey_median(utilities_ratio, na.rm = TRUE),
      
      mean_food_ratio = survey_mean(food_ratio, na.rm = TRUE),
      median_food_ratio = survey_median(food_ratio, na.rm = TRUE),
      
      mean_essential_ratio = survey_mean(essential_ratio, na.rm = TRUE),
      median_essential_ratio = survey_median(essential_ratio, na.rm = TRUE),
      
      mean_Housing.Utilities_ratio = survey_mean(Housing.Utilities_ratio, na.rm = TRUE),
      median_Housing.Utilities_ratio = survey_median(Housing.Utilities_ratio, na.rm = TRUE)
      
  ) %>%
    select(-(contains("_se"))) %>%
  pivot_longer(cols = starts_with("mean_") | starts_with("median_"),
               names_to = c(".value", "measure"),
               names_sep = "_") %>%
  tibble()

```

# 8 What share of all income (work + benefits) is taken up by housing + energy for people renting privately, in social housing & mortgage holders.
```{r}
#What share of all income (work + benefits) is taken up by housing + energy for people renting privately, in social housing & mortgage holders.
benefits_pop<-bm %>% filter(Income.Total.Benefits>0) 

variables_to_group_by <- c("Derived.Household.Type.LCFS", 
                           "Housing",
                           "Role.Type",
                           "Disabled", 
                           "UC",
                           "ethnicity.HRP"
                           )

driver_benefits<-benefits_pop %>% 
  mutate(housing_ratio = case_when(Expenditure.Housing <= 0 ~ NA,
                                   TRUE ~ Expenditure.Housing / Income.Total),
         rent_ratio = case_when(Expenditure.Rent <= 0 ~ NA,
                                   TRUE ~ Expenditure.Rent / Income.Total),
         mortgage_ratio = case_when(Expenditure.Mortgage <= 0 ~ NA,
                                   TRUE ~ Expenditure.Mortgage / Income.Total),
         transport_ratio = case_when(Expenditure.Transport <= 0 ~ NA,
                                   TRUE ~ Expenditure.Transport / Income.Total),
         utilities_ratio = case_when(Expenditure.Utilities.MART <= 0 ~ NA,
                                   TRUE ~ Expenditure.Utilities.MART / Income.Total),
         food_ratio = case_when(Expenditure.Food.House.Keeping.MART <= 0 ~ NA,
                                   TRUE ~ Expenditure.Food.House.Keeping.MART / Income.Total),
         essential_ratio = case_when(Expenditure.Total <= 0 ~ NA,
                                   TRUE ~ Expenditure.Total / Income.Total)
  )
        
result_driver_utilities_benefits <- map_dfr(variables_to_group_by, function(var) {
  driver_benefits %>% 
as_survey_design(weights = Weight.LCFS) %>%
    group_by(Financial.Year, !!sym(paste0("Demographic.", var))) %>%
    summarise(
      mean_utilities_ratio = survey_mean(utilities_ratio, na.rm = TRUE),
      median_utilities_ratio = survey_median(utilities_ratio, na.rm = TRUE)
    ) %>%
    select(-(contains("_se"))) %>%
    tibble()
})


result_driver_housing_benefits <- map_dfr(variables_to_group_by, function(var) {
  driver_benefits %>% 
as_survey_design(weights = Weight.LCFS) %>%
    group_by(Financial.Year, !!sym(paste0("Demographic.", var))) %>%
    summarise(
      mean_housing_ratio = survey_mean(housing_ratio, na.rm = TRUE),
      median_housing_ratio = survey_median(housing_ratio, na.rm = TRUE)
    )  %>%
    select(-(contains("_se"))) %>%
    tibble()
})



result_driver_utilities_benefits <- map_dfr(variables_to_group_by, function(var) {
  driver_benefits %>% 
as_survey_design(weights = Weight.LCFS) %>%
    group_by(Financial.Year, !!sym(paste0("Demographic.", var))) %>%
    summarise(
      mean_food_ratio = survey_mean(food_ratio, na.rm = TRUE),
      median_food_ratio = survey_median(food_ratio, na.rm = TRUE)
    )  %>%
    select(-(contains("_se"))) %>%
    tibble()
})

```


```{r}
calculate_all_ratios <- function(data, weights, filter_condition, ratio_names) {
  data %>%
    filter(filter_condition) %>%
    mutate_at(ratio_names, list(~ case_when(. <= 0 ~ NA, TRUE ~ ./Income.Total))) %>%
    as_survey_design(weights = Weight.LCFS)
}

variables_to_group_by <- c(
  "Derived.Household.Type.LCFS",
  "Housing",
  "Role.Type",
  "Disabled",
  "UC",
  "ethnicity.HRP"
)

benefits_pop <- bm %>% filter(Income.Total.Benefits > 0)

ratio_names <- c("Expenditure.Housing", 
                 "Expenditure.Utilities.MART", 
                 "Expenditure.Food.House.Keeping.MART")

result_driver_all_ratios <- map_dfr(variables_to_group_by, function(var) {
  calculate_all_ratios(benefits_pop, Weight.LCFS, TRUE, ratio_names) %>%
    group_by(Financial.Year, !!sym(paste0("Demographic.", var))) %>%
    summarise(across(all_of(ratio_names),
                     list(mean_ratio = ~ survey_mean(., na.rm = TRUE),
                          median_ratio = ~ survey_median(., na.rm = TRUE)),
                     .names = "{.col}_{.fn}")) %>%
    mutate(across(contains("_mean_ratio"), ~ case_when(. <= 0 ~ NA, TRUE ~ .))) %>%
    select(-contains("_se")) %>%
    tibble()
})

write.csv(result_driver_all_ratios, "result_driver_all_ratios.csv")
```

# To get to know what is the weekly expenditures 
```{r}
expenditure_overall<-bm %>% 
  as_survey_design(weights = Weight.LCFS) %>%
  filter(Financial.Year == "2023-24") %>% 
  mutate(across(starts_with("Expenditure"), ~ ifelse(. == 0, NA, .))) %>%
  summarise(
    across(starts_with("Expenditure"), list(mean = ~ survey_mean(., na.rm = TRUE), 
                                             median = ~ survey_median(., na.rm = TRUE)),
                                             .names = "{.col}_{.fn}")
  ) %>%
  select(-(contains("_se"))) %>%
  pivot_longer(cols = starts_with("Expenditure"),
               names_to = c("variable", ".value"),
               names_sep = "_") %>%
  tibble()
write.csv(expenditure_overall, "expenditure_overall.csv")



```

